---
Title: Geological R project
output: github_document
---

```{r}
knitr::opts_chunk$set(
  warning = FALSE, # show warnings
  message = TRUE, # show messages
  error = TRUE, # do not interrupt generation in case of errors,
  echo = TRUE  # show R code
)
```

# Geology masters code

The aim of this folder is to.... statistically and graphically observe the differences in a variety of trace element concentrations in different forms of Carrollite in the central African Copperbelt, then a principle component analysis on the different styles of mineralization

load packages in R 

```{r}
library(ggplot2) 
pacman::p_load(tidyverse)
library(wesanderson)
library(FactoMineR)
library(factoextra)

```

The use of ggplots2 to create the box and whisker
The grouped plots that will be used to test the geological data

loading the excel data via clipboard
this will not work without it being copied!!

```{r}
# This is just an easy way of copying in the data and then once local storage is required the write.csv can 
# be used so that the data can be knitted
#Car8 <- read.table(pipe("pbpaste"), sep="\t", header = TRUE)
# moving it to csv files for the knit function
#write.csv(Car8,"/Users/charlesrandell/Other/Random projects/Box and Whisker/data/Car8.csv", row.names = FALSE)

setwd("/Users/charlesrandell/Other/Random projects/Box and Whisker/data")

CarStrat <- read.csv("CarStrat.csv", header = TRUE)

CarDissem <- read.csv("CarDissem.csv", header = TRUE)

CarCross <- read.csv("CarCross.csv", header = TRUE)

CarJack <- read.csv("CarJack.csv", header = TRUE)

combination <- read.csv("Combination.csv", header = TRUE)

Chal88 <- read.csv("Chal88.csv", header = TRUE)

Chal8 <- read.csv("Chal8.csv", header = TRUE)

Car88 <- read.csv("Car8.csv", header = TRUE)

Car8 <- read.csv("Car88.csv", header = TRUE)

```

## Tidyr
Lets see what happens...
Changing the data imported in wide format into a long tidy format
```{r}

TidyCarCross <- CarCross %>% tidyr::gather(Header, val)

TidyCarDissem <- CarDissem %>% tidyr::gather(Header, val)

TidyCarStrat <- CarStrat  %>% tidyr::gather(Header, val)

TidyCarJack <- CarJack  %>% tidyr::gather(Header, val)

```

## dplyr
This creates the new variable that has the extra column with the added indicator to be merged to the final dataset.
The use of the tidyverse to convert the data into a long format rather that a wide so that ggplot2 can create the boxplot with it

```{r}
Tcarcross <- TidyCarCross %>% mutate(Classes = "4CarCross")

Tcardissem <- TidyCarDissem %>% mutate(Classes = "2CarDissem")

Tcarstrat <- TidyCarStrat %>% mutate(Classes = "1CarStrat")

Tcarjack <- TidyCarJack %>% mutate(Classes = "3CarJack")


```

## binding and filtering the data
Bind the datasets together
Secondly the readings that are below the detectable level are filtered out 
```{r}
Final <- rbind.data.frame(Tcarstrat, Tcardissem, Tcarcross, Tcarjack)
Final <- Final %>% filter(val != "Below LOD")
Final <- Final %>% filter(Header != "Te_ppm_m125")
```

## gsub to format elements
mutate function that can remove certain phrases within values in a column 

```{r}
Final <- Final %>% mutate(Header = gsub("_ppm", "", Header))
Final <- Final %>% mutate(Header = gsub("_m126", "", Header))
```

Convert the variable Classes and Header from a character to a factor variable
Convert the value variable from a character to a numerical variable - this is very important and the code will not run without it
```{r}
Final[, 'Classes'] <- as.factor(Final[, 'Classes'])
Final[, 'val'] <- as.numeric(Final[, 'val'])
Final[, 'Header'] <- as.factor(Final[, 'Header'])
```


Firstly the outliers are removed, any value that exceeds 150 is taked out
```{r}
Final <- Final %>% filter(val < 150)
```

## ggplot to create graphs 
The plot is constructed using ggplot, with two iterations of the same graph, the first shows the raw output on a linear scale
And second has a log scale y axis, along with updated labels and different colour palette
```{r}
# The Main ggplot

sfplot <- ggplot(Final, aes(x=Header, y=val, fill=Classes)) + 
    geom_boxplot(outlier.size = 0.01) + stat_summary(fun = mean, shape = 4, aes(group=Classes), position=position_dodge(0.75), show.legend = FALSE, color="black", size=0.25)

sfplot

# For some reason this is what worked for the color codes to range from 2-5

wes1 <- wes_palette("FantasticFox1")[2:5]
roy2 <- wes_palette("Royal2")[1:3]


# The Final Wes Anderson themed plot

sfplot  +
coord_trans(y = "log") + scale_y_continuous(breaks = c(0.01, 0.1, 1, 10, 100), labels = c(0.01, 0.1, 1, 10, 100)) + labs(x="Elements", y = "Log (Concentrations) / ppm") + theme(legend.position="bottom") + scale_fill_manual(guide = guide_legend(title = ""), breaks=c("1CarStrat", "2CarDissem", "3CarJack", "4CarCross"), values=c(wes1), labels=c("Stratiform Carrollite", "Disseminated Carrollite", "Jack Vein Carrollite", "Cross-Cutting Vein Carrollite")) + theme(axis.text.x = element_text(size=12, face = "bold"), axis.text.y = element_text(size=12, face = "bold"), legend.text=element_text(size=14), axis.title=element_text(size=15,face="bold"))

```


## t tests

```{r}
CarCross <- mutate_all(CarCross, function(x) as.numeric(as.character(x)))
CarDissem <- mutate_all(CarDissem, function(x) as.numeric(as.character(x)))
CarJack <- mutate_all(CarJack, function(x) as.numeric(as.character(x)))
CarStrat <- mutate_all(CarStrat, function(x) as.numeric(as.character(x)))


myfunc <-function(x) { 
  x %>%
    mutate_all(function(x) as.numeric(as.character(x)))
}

# apply function to make all numeric

Car8 <- myfunc(Car8)
Car88 <- myfunc(Car88)
Chal8 <- myfunc(Chal8)
Chal88 <- myfunc(Chal88)


# make the nas zero

Car8[is.na(Car8)] <- 0
Car88[is.na(Car88)] <- 0
Chal8[is.na(Chal8)] <- 0
Chal88[is.na(Chal88)] <- 0

# apply function to all datasets 

combination <- mutate_all(combination, function(x) as.numeric(as.character(x)))

CarCross[is.na(CarCross)] <- 0
combination[is.na(combination)] <- 0

t.test(CarCross$Zn_ppm, combination$Zn_ppm)
t.test(CarCross$Te_ppm, combination$Te_ppm)
t.test(CarCross$Sn_ppm, combination$Sn_ppm)
t.test(CarCross$Se_ppm, combination$Se_ppm)
t.test(CarCross$Sb_ppm, combination$Sb_ppm)
t.test(CarCross$Pb_ppm, combination$Pb_ppm)
t.test(CarCross$Mo_ppm, combination$Mo_ppm)
t.test(CarCross$In_ppm, combination$In_ppm)
t.test(CarCross$Ge_ppm, combination$Ge_ppm)
t.test(CarCross$Ga_ppm, combination$Ga_ppm)
t.test(CarCross$Cd_ppm, combination$Cd_ppm)
t.test(CarCross$Bi_ppm, combination$Bi_ppm)
t.test(CarCross$Au_ppm, combination$Au_ppm)
t.test(CarCross$Ag_ppm, combination$Ag_ppm)

confidence_interval <- function(vector, interval) {
  # Standard deviation of sample
  vec_sd <- sd(vector)
  # Sample size
  n <- length(vector)
  # Mean of sample
  vec_mean <- mean(vector)
  # Error according to t distribution
  error <- qt((interval + 1)/2, df = n - 1) * vec_sd / sqrt(n)
  # Confidence interval as a vector
  result <- c("lower" = vec_mean - error, "upper" = vec_mean + error)
  return(result)
}
```

## t test for second dataset Car

```{r}
t.test(Car8$Zn_ppm, Car88$Zn_ppm)
t.test(Car8$Te_ppm, Car88$Te_ppm)
t.test(Car8$Sn_ppm, Car88$Sn_ppm)
t.test(Car8$Se_ppm, Car88$Se_ppm)
t.test(Car8$Sb_ppm, Car88$Sb_ppm)
t.test(Car8$Pb_ppm, Car88$Pb_ppm)
t.test(Car8$Mo_ppm, Car88$Mo_ppm)
t.test(Car8$In_ppm, Car88$In_ppm)
t.test(Car8$Ge_ppm, Car88$Ge_ppm)
t.test(Car8$Ga_ppm, Car88$Ga_ppm)
t.test(Car8$Cd_ppm, Car88$Cd_ppm)
t.test(Car8$Bi_ppm, Car88$Bi_ppm)
t.test(Car8$Au_ppm, Car88$Au_ppm)
t.test(Car8$Ag_ppm, Car88$Ag_ppm)
```


## t test for third dataset Chal

```{r}

t.test(Chal8$Zn_ppm, Chal88$Zn_ppm)
t.test(Chal8$Te_ppm, Chal88$Te_ppm)
t.test(Chal8$Sn_ppm, Chal88$Sn_ppm)
t.test(Chal8$Se_ppm, Chal88$Se_ppm)
t.test(Chal8$Sb_ppm, Chal88$Sb_ppm)
t.test(Chal8$Pb_ppm, Chal88$Pb_ppm)
t.test(Chal8$Mo_ppm, Chal88$Mo_ppm)
t.test(Chal8$In_ppm, Chal88$In_ppm)
t.test(Chal8$Ge_ppm, Chal88$Ge_ppm)
t.test(Chal8$Ga_ppm, Chal88$Ga_ppm)
t.test(Chal8$Cd_ppm, Chal88$Cd_ppm)
t.test(Chal8$Bi_ppm, Chal88$Bi_ppm)
t.test(Chal8$Au_ppm, Chal88$Au_ppm)
t.test(Chal8$Ag_ppm, Chal88$Ag_ppm)

```

# pca

This section will perform a principle component analysis on the all of the grouped data with the "Below LOD" 
reduced to zero, and an identification style attached to each dataset. 

```{r}
CarStrat$Classes <- 'CarStrat'
CarCross$Classes <- "CarCross"
CarDissem$Classes <- "CarDissem"
CarJack$Classes <- "CarJack"

pcat <- rbind(CarStrat, CarCross, CarDissem, CarJack)
pcat[, 'Classes'] <- as.factor(pcat[, 'Classes'])

car.pca <- PCA(pcat %>% select(-Classes), graph = FALSE)

pcat[is.na(pcat)] <- 0

fviz_contrib(car.pca, choice = "var", axes = 1)


fviz_pca_biplot(car.pca, habillage = pcat$Classes, addEllipses = TRUE, 
    col.var = "red", alpha.var = "cos2", label = "var") + scale_color_brewer(palette = "Dark2") + 
    theme_minimal()

```


