---
output: github_document
  md_document:
    variant: markdown_github
---

# Purpose

The aim of this folder is to.... create a grouped box and whisker plot with the ggplot2 package that shows a variety of trace element concentrations in different forms of Carrollite in the central African Copperbelt 

## load packages in R 

```{r}
library(ggplot2) 
pacman::p_load(tidyverse)
library(wesanderson)

```

#### The use of ggplots2 to create the box and whisker
## The grouped plots that will be used to test the geological data

### loading the excel data via clipboard
### this will not work without it being copied!!

```{r}
CarStrat <- read.table(pipe("pbpaste"), sep="\t", header = TRUE)

CarDissem <- read.table(pipe("pbpaste"), sep="\t", header = TRUE)

CarCross <- read.table(pipe("pbpaste"), sep="\t", header = TRUE)

CarJack <- read.table(pipe("pbpaste"), sep="\t", header = TRUE)

```


## Lets see what happens...
##
```{r}

TidyCarCross <- CarCross %>% tidyr::gather(Header, val)

TidyCarDissem <- CarDissem %>% tidyr::gather(Header, val)

TidyCarStrat <- CarStrat  %>% tidyr::gather(Header, val)

TidyCarJack <- CarJack  %>% tidyr::gather(Header, val)

```


## This creates the new variable that has the extra column with the added indicator to be merged to the final dataset
## The use of the tidyverse to convert the data into a long format rather that a wide so that ggplot2 can create the boxplot with it

```{r}
Tcarcross <- TidyCarCross %>% mutate(Classes = "4CarCross")

Tcardissem <- TidyCarDissem %>% mutate(Classes = "2CarDissem")

Tcarstrat <- TidyCarStrat %>% mutate(Classes = "1CarStrat")

Tcarjack <- TidyCarJack %>% mutate(Classes = "3CarJack")


```


### Bind the datasets together
### Secondly the readings that are below the detectable level are filtered out 
```{r}
Final <- rbind.data.frame(Tcarstrat, Tcardissem, Tcarcross, Tcarjack)
Final <- Final %>% filter(val != "Below LOD")
Final <- Final %>% filter(Header != "Te_ppm_m125")
```


## mutate function that can remove certain phrases within values in a column 

```{r}
Final <- Final %>% mutate(Header = gsub("_ppm", "", Header))
Final <- Final %>% mutate(Header = gsub("_m126", "", Header))
```



### Convert the variable Classes and Header from a character to a factor variable
### Convert the value variable from a character to a numerical variable - this is very important and the code will not run without it

```{r}
Final[, 'Classes'] <- as.factor(Final[, 'Classes'])
Final[, 'val'] <- as.numeric(Final[, 'val'])
Final[, 'Header'] <- as.factor(Final[, 'Header'])
```


## Firstly the outliers are removed, any value that exceeds 150 is taked out

```{r}
Final <- Final %>% filter(val < 150)
```


## The plot is constructed using ggplot, with two iterations of the same graph, the first shows the raw output on a linear scale
## And second has a log scale y axis, along with updated labels and different colour palette
```{r}
## The Main ggplot
sfplot <- ggplot(Final, aes(x=Header, y=val, fill=Classes)) + 
    geom_boxplot(outlier.size = 0.01) + stat_summary(fun = mean, shape = 4, aes(group=Classes), position=position_dodge(0.75), show.legend = FALSE, color="black", size=0.25)

sfplot
## For some reason this is what worked for the color codes to range from 2-5

wes1 <- wes_palette("FantasticFox1")[2:5]
roy2 <- wes_palette("Royal2")[1:3]
c1 <- colorspaces("#99FF99")
## the first iteration

sfplot + scale_x_discrete(limits=c("Zn_ppm", "Ga_ppm", "Ge_ppm", "Se_ppm", "Mo_ppm", "Ag_ppm", "Cd_ppm", "In_ppm", "Sb_ppm", "Te_ppm_m125", "Te_ppm_m126", "Au_ppm", "Pb_ppm", "Bi_ppm")) 


## The Final Wes Anderson themed plot

sfplot  +
coord_trans(y = "log") + scale_y_continuous(breaks = c(0.01, 0.1, 1, 10, 100), labels = c(0.01, 0.1, 1, 10, 100)) + labs(x="Elements", y = "Log (Concentrations) / ppm") + theme(legend.position="bottom") + scale_fill_manual(guide = guide_legend(title = ""), breaks=c("1CarStrat", "2CarDissem", "3CarJack", "4CarCross"), values=c(wes1), labels=c("Stratiform Carrollite", "Disseminated Carrollite", "Jack Vein Carrollite", "Cross-Cutting Vein Carrollite")) + theme(axis.text.x = element_text(size=12, face = "bold"), axis.text.y = element_text(size=12, face = "bold"), legend.text=element_text(size=14), axis.title=element_text(size=15,face="bold"))

```






